# Reproducible Research: Peer Assessment 1

## Loading and preprocessing the data

* set working directory 
* laod data
* aggregate steps per day

```{r}
setwd("~/Documents/SWE/udacity_coursera_edx/datasciencecoursera/RepData_PeerAssessment1/")
activity <- read.csv("activity.csv")

steps_per_day <- aggregate(activity$steps, by=list(activity$date), sum)
names(steps_per_day) <- c("date", "steps")
```


## What is mean total number of steps taken per day?

```{r fig.width=10, fig.height=6}
hist(steps_per_day$steps, xlab="Steps per Day", main="Histogram of Steps per Day")
```

```{r}
mean_steps <- mean(steps_per_day$steps, na.rm = TRUE)
median_steps <- median(steps_per_day$steps, na.rm = TRUE)
options(scipen=1, digits=1)
```

- **Mean** total number of steps taken per day: **`r mean_steps`**
- **Median** total number of steps taken per day: **`r median_steps`**


## What is the average daily activity pattern?

```{r}
steps_per_interval <- aggregate(activity$steps, by=list(activity$interval), mean, na.rm = TRUE)
names(steps_per_interval) <- c("interval", "steps")
```

```{r fig.width=10, fig.height=6}
plot(x=steps_per_interval$interval, y=steps_per_interval$steps, type="l",
     xlab="Interval ID", ylab="Steps", main="Daily Activity")
```

```{r}
top_id <- steps_per_interval$interval[steps_per_interval$steps==max(steps_per_interval$steps)]
```
The **Top ID**, i.e. theID of the 5-minute interval, which on average across all the days in the dataset, contains the maximum number of steps is **`r top_id`**


## Imputing missing values

TODO


## Are there differences in activity patterns between weekdays and weekends?

TODO


<hr>
<a href="#top">top</a>

